---
title: "mverse"
subtitle: "An R library for Teaching and Conducting<br/>Multiverse Analysis"
author: "Michael Jongho Moon"
institute: "PhD Student, DoSS, University of Toronto"
date: "SSC 2021 Annual Meeting, June 7, 2021"
output:
  xaringan::moon_reader:
    css: ["style/sansation/stylesheet.css", "style/theme.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      titleSlideClass: [middle, left]
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggtext)
library(igraph)
library(ggraph)
library(mverse)
hdata <- hurricane %>%
  mutate(
    nameyear = paste(Name, Year, sep = ", ")
  ) %>%
  rename(
    name = Name,
    deaths = alldeaths,
    femininity = MasFem,
    year = Year,
    min_pressure = Minpressure_Updated_2014,
    max_wind_speed = HighestWindSpeed,
    damage = NDAM,
  ) %>%
  select(
    name, nameyear, deaths, femininity, year,
    min_pressure, max_wind_speed, damage
  )
```

.nav-left[# Collaborators]

.main-right[
.third[
## .accent2[`mverse` Education Development]
.accent2[
Alison Gibbs, PhD  
Nathan Taback, PhD  
Fanny Chevalier, PhD  

University of Toronto]
]
.third[
## .accent3[`mverse` Package Development]
.accent3[
Mingwei Xu  
Haoda Li  

University of Toronto]
]
.third.accent1[
## .accent1[`multiverse` Package Development]
.accent1[
Abhraneel Sarma  
Matthew Kay, PhD

Northwestern University]
]]

---

layout:true

# Example: Hurricane Names and Gender-based Expectations

---

```{r}
topdeaths <- hdata %>%
  mutate(nameyear = factor(nameyear, levels = nameyear[order(deaths)])) %>%
  arrange(-deaths) %>%
  slice(1:10)
topdamages <- hdata %>%
  mutate(nameyear = factor(nameyear, levels = nameyear[order(damage)])) %>%
  arrange(-damage) %>%
  slice(1:10)
```

.nav-right[

### Do people underestimate the severity of a hurricane when it's named with a female name? 

]

.main-right[.half[

```{r out.width="90%", fig.asp=1, fig.align="center", dpi=144}
p1 <- topdeaths %>%
  ggplot(aes(x = nameyear, y = deaths, fill = femininity > 6)) +
  theme_void() +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "10 deadliest Atlantic hurricanes",  
       subtitle = "in the U.S between 1950 and 2012.") +
  scale_y_continuous(labels = scales::comma, name = "Fatalities", position = "right") +
  scale_fill_manual(values = c(8, 1)) +
  coord_flip() +
  annotate("richtext", x = 4, y = max(hdata$deaths), hjust = 1, vjust = 0, 
           fill = NA, label.color = NA, color = "#696969", size = 8,
           label = paste0(
             "<b style='color:black'>", topdeaths %>% filter(femininity > 6) %>% nrow(), 
             "</b> out of the 10<br />deadliest hurricanes<br />had <i>female</i> names."
            )) +
  geom_text(aes(label = name), color = "white", size = 6, hjust = 1, nudge_y = -25,
            data = topdeaths %>% filter(femininity > 6, deaths > 350)) +
  geom_text(aes(label = name), color = "black", size = 6, hjust = 0, nudge_y = 25,
            data = topdeaths %>% filter(femininity > 6, deaths <= 350)) +
  geom_text(aes(label = name), color = 8, size = 6, hjust = 0, nudge_y = 25,
            data = topdeaths %>% filter(femininity <= 6)) +
  theme(
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 20),
    axis.text.x.top = element_text(size = 14),
    axis.title.x.top = element_text(size = 16, hjust = 1)
  )
p1
```

]

.half[

```{r out.width="90%", fig.asp=1, fig.align="center", dpi=144}
p2 <- topdamages %>%
  ggplot(aes(x = nameyear, y = damage, fill = femininity > 6)) +
  theme_void() +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "10 mosting damaging Atlantic hurricanes",
       subtitle = "in the U.S between 1950 and 2012.",
       caption = "1. Financial damage amount in USD normlized to 2013 values.") +
  scale_y_continuous(labels = scales::comma, position = "right",
                     name = expression(Damage^1)) +
  scale_fill_manual(values = c(8, 1)) +
  coord_flip() +
  annotate("richtext", x = 2, y = max(hdata$damage), hjust = 1, vjust = 0, 
           fill = NA, label.color = NA, color = "#696969", size = 8,
           label = paste0(
             "<b style='color:black'>", topdamages %>% filter(femininity > 6) %>% nrow(), 
             "</b> out of the 10<br />most damaging hurricanes<br />had <i>female</i> names."
            )) +
  geom_text(aes(label = name), color = "white", size = 6, hjust = 1, nudge_y = -1000,
            data = topdamages %>% filter(femininity > 6)) +
  geom_text(aes(label = name), color = "black", size = 6, hjust = 1, nudge_y = -1000,
            data = topdamages %>% filter(femininity <= 6)) +
  theme(
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 20),
    plot.caption = element_text(size = 14),
    axis.text.x.top = element_text(size = 14),
    axis.title.x.top = element_text(size = 16, hjust = 1),
    plot.margin = margin(r = 10)
  )
p2
```

]]

.footnote[
Original data (n=94) retrieved from Jung, K., Shavitt, S., Viswanathan, M., and Hilbe, J.M. (2014) _Female hurricanes are deadlier than male hurricanes._  
Proceedings of the National Academy of Sciences, 111(24), 8782-8787, https://doi.org/10.1073/pnas.1402786111
]

---

.nav-right[

### Do people underestimate the severity of a hurricane when it's named with a female name? 

.selected[Did hurricanes with *feminine* names lead to __more fatalities given equal strength__ because their names didn't motivate as much preparedness as hurricanes with *masculine* names?]

]

.main-right[.half[

```{r out.width="90%", fig.asp=1, fig.align="center", dpi=144}
p1
```

]

.half[

```{r out.width="90%", fig.asp=1, fig.align="center", dpi=144}
p2
```

]]

.footnote[
Original data (n=94) retrieved from Jung, K., Shavitt, S., Viswanathan, M., and Hilbe, J.M. (2014) _Female hurricanes are deadlier than male hurricanes._  
Proceedings of the National Academy of Sciences, 111(24), 8782-8787, https://doi.org/10.1073/pnas.1402786111
]

---

.nav-right[

### Do people underestimate the severity of a hurricane when it's named with a female name? 


Consider a data set with the following information:

.small[
+   Femininity rating on hurricane names (1:very masculine; 11: very feminine)
+   Total damage dollar amount normalized to 2013 USD
+   Total fatality counts
+   Maximum wind speed
+   Minimum pressure
+   Year
]
]

.main-right[.half[

```{r out.width="90%", fig.asp=1, fig.align="center", dpi=144}
p1
```

]

.half[

```{r out.width="90%", fig.asp=1, fig.align="center", dpi=144}
p2
```

]]

.footnote[
Original data (n=94) retrieved from Jung, K., Shavitt, S., Viswanathan, M., and Hilbe, J.M. (2014) _Female hurricanes are deadlier than male hurricanes._  
Proceedings of the National Academy of Sciences, 111(24), 8782-8787, https://doi.org/10.1073/pnas.1402786111
]

---

```{r}
twodeadliest <- hdata %>% arrange(-deaths) %>% slice(1:2)
ann <- paste0("<b style='color:#DF536B;'>",
  (twodeadliest %>% pull(name))[1],
  "</b> was by far the deadliest hurricane<br/>causing <b style='color:#DF536B;'>",
  (twodeadliest %>% pull(deaths))[1] %>% format(big.mark = ","),
  "</b> deaths in ", (twodeadliest %>% pull(year))[1], ".",
  "<br/>The second deadliest hurricane during the period,<br/><b style='color:#2297E6;'>",
  (twodeadliest %>% pull(name))[2],
  "</b>, resulted in <b style='color:#2297E6;'>", 
  (twodeadliest %>% pull(deaths))[2],
  "</b> deaths in ", (twodeadliest %>% pull(year))[2],
  "."
)
```

.nav-right[

Should we remove _Katrina_ or _Katrina_ and _Audrey_ from data based on the distribution of fatalities?

]

.footnote[
Original data (n=94) retrieved from Jung, K., Shavitt, S., Viswanathan, M., and Hilbe, J.M. (2014) _Female hurricanes are deadlier than male hurricanes._  
Proceedings of the National Academy of Sciences, 111(24), 8782-8787, https://doi.org/10.1073/pnas.1402786111
]


.main-right[

```{r out.width="90%", fig.asp=1/2, fig.align="center", dpi=144}
hdata %>%
  ggplot(aes(x = deaths, fill = factor(if_else(
    name == (twodeadliest %>% pull(name))[1], 1, 
    if_else(name == (twodeadliest %>% pull(name))[2], 2, 0))))) +
  theme_void() +
  geom_dotplot(binwidth = 20, color = "white", dotsize = 0.8, stackratio = 0.75, 
               show.legend = FALSE) +
  annotate("richtext", x = 200, y = 0.8, 
           hjust = 0, vjust = 1, 
           fill = NA, label.color = NA, color = "#696969", size = 4, 
           label = ann) +
  geom_text(aes(label = name, y = 0.1), data = twodeadliest, color = c(2, 4)) +
  scale_fill_manual(values = c(8, 2, 4)) +
  labs(x = "Fatalities") +
  theme(
    plot.title = element_text(size = 24),
    plot.subtitle = element_text(size = 20),
    plot.caption = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.title.x = element_text(size = 16, hjust = 1),
    axis.line.x = element_line()
  )
```

]

---

.half[

### Michael

<b>1.</b> Remove _Katrina_ and _Audrey_.

]

.half[

### Mary

<b>1.</b> Remove _Katrina_ only.

]

<hr/>

<b>2.</b> Fit a regression model with _Femininity_, _Damage_, and their interaction.

$$\text{Fatalities}_i\sim \text{Pois}\left(\mu_i\right)$$
$$\log\left(\mu_i\right) = \beta_0 + \beta_1\text{Damage}_i + \beta_2\text{Femininity}_i + \beta_3\text{Damage}_i\times\text{Femininity}_i$$


<hr/>

<b>3.</b> Make an inference based on the interaction effect $\beta_3$.

---

```{r}
data_1 <- hdata %>%
  filter(! name %in% c("Katrina", "Audrey"))
data_2 <- hdata %>%
  filter(! name %in% c("Katrina"))
fit_1 <- glm(deaths ~ damage * femininity, data = data_1, family = "poisson")
fit_2 <- glm(deaths ~ damage * femininity, data = data_2, family = "poisson")
ilink_1 <- family(fit_1)$linkinv
ilink_2 <- family(fit_2)$linkinv
est_1 <- summary(fit_1)$coef
est_2 <- summary(fit_2)$coef
```

.nav-right[

**Michael**: stronger hurricanes with a feminine name are under estimated

]

.main-right[

```{r fig.align="center", fig.asp=9/16, out.width="80%", dpi=144}
as_tibble(rbind(est_1[4,], est_2[4,])) %>%
  mutate(
    model = factor(c("Michael", "Mary"), levels = c("Michael", "Mary")),
    ratio = ilink_1(Estimate),
    lwr = ilink_1(Estimate - qnorm(0.975) * `Std. Error`),
    upr = ilink_1(Estimate + qnorm(0.975) * `Std. Error`)
  ) %>%
  ggplot(aes(y = ratio, ymin = lwr, ymax = upr, x = model, color = model)) +
  theme_void() +
  geom_pointrange(show.legend = FALSE)  +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_color_manual(values = c(1, "white")) +
  scale_y_continuous(breaks = 1) +
  scale_x_discrete(position = "top", labels = c("Michael", "")) +
  labs(y = "Interaction Effect (Ratio)", 
       title = "Estimated Interaction Effect and 95% CI\nof Damage (2013 USD) x Femininity (1-11 scale)") +
  theme(
    plot.title = element_text(size = 18, margin = margin(b = 16)),
    plot.title.position = "plot",
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    axis.title.y = element_text(size = 16, angle = 90, hjust = 1)
  )
```

]

---

.nav-right[


**Michael**: stronger hurricanes with a feminine name are under estimated

**Mary**: there is no evidence of gender-based expectations for hurricanes

]

.main-right[

```{r fig.align="center", fig.asp=9/16, out.width="80%", dpi=144}
p3 <- as_tibble(rbind(est_1[4,], est_2[4,])) %>%
  mutate(
    model = factor(c("Michael", "Mary"), levels = c("Michael", "Mary")),
    ratio = ilink_1(Estimate),
    lwr = ilink_1(Estimate - qnorm(0.975) * `Std. Error`),
    upr = ilink_1(Estimate + qnorm(0.975) * `Std. Error`)
  ) %>%
  ggplot(aes(y = ratio, ymin = lwr, ymax = upr, x = model)) +
  theme_void() +
  geom_pointrange()  +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_y_continuous(breaks = 1) +
  scale_x_discrete(position = "top") +
  labs(y = "Interaction Effect (Ratio)", 
       title = "Estimated Interaction Effect and 95% CI\nof Damage (2013 USD) x Femininity (1-11 scale)") +
  theme(
    plot.title = element_text(size = 18, margin = margin(b = 16)),
    plot.title.position = "plot",
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    axis.title.y = element_text(size = 16, angle = 90, hjust = 1)
  )
p3
```

]

---

.nav-right[


**Michael**: stronger hurricanes with a feminine name are under estimated

**Mary**: there is no evidence of gender-based expectations for hurricanes

# .selected[_Which result should they report?_]

]

.main-right[

```{r fig.align="center", fig.asp=9/16, out.width="80%", dpi=144}
p3
```

]

---
layout:false

.nav-left[

# Multiverse Analysis (Steegen et al. 2016)<sup>1</sup>

+   Performing and communicating all alternative analyses based on reasonable alternative paths
+   Improves **transparency** and **robustness** of the reported findings

]

.footnote[
1. Steegen, S., Tuerlinckx, F., Gelman, A., and Vanpaemel, W. (2016) _Increasing transparency through a multiverse analysis._ Perspectives on Psychological Science, 11(5), 702-712,  https://doi.org/10.1177/1745691616658637
]

--


## Example: A Multiverse of Hurricane Name Analaysis

```{r}
kable(tibble(`Analysis Decisions` = c(
  "> Exclude 2 outliers with most fatalities",
  "> Exclude 1 outlier with most fatalities",
  "> Exclude 2 outliers with most fatalities, 2 with greatest damages",
  "> Normalized damage amount",
  "> Maximum wind speed",
  "> Minimum pressure",
  "> Femininity x hurricane strength interaction",
  "> Femininity main effect")),
  col.names = c("")) %>%
  pack_rows("Which hurricanes should we analyse?", 1, 3) %>%
  pack_rows("Which variable to use to capture hurricane strengths?", 4, 6) %>%
  pack_rows("Which effect should the inference be based on?", 7, 8)
```

---

.nav-left[

# Multiverse Analysis (Steegen et al. 2016)<sup>1</sup>

+   Performing and communicating all alternative analyses based on reasonable alternative paths
+   Improves **transparency** and **robustness** of the reported findings

]


```{r out.width="45%", warning=FALSE, message=FALSE, dpi=144}
edges <- as_tibble(rbind(
  c("0", "Exclude 2 deadliest"),
  c("0", "Exclude 1 deadliest"),
  c("0", "Exclude 1 deadliest and 1 most damaging"),
  c("Exclude 2 deadliest", "Normalized damage amount - 1"),
  c("Exclude 2 deadliest", "Maximum wind speed - 1"),
  c("Exclude 2 deadliest", "Minimum pressure - 1"),
  c("Exclude 1 deadliest", "Normalized damage amount - 2"),
  c("Exclude 1 deadliest", "Maximum wind speed - 2"),
  c("Exclude 1 deadliest", "Minimum pressure - 2"),
  c("Exclude 1 deadliest and 1 most damaging", "Normalized damage amount - 3"),
  c("Exclude 1 deadliest and 1 most damaging", "Maximum wind speed - 3"),
  c("Exclude 1 deadliest and 1 most damaging", "Minimum pressure - 3"),
  c("Normalized damage amount - 1", "Interaction - 1 - 1"),
  c("Normalized damage amount - 1", "Main effect - 1 - 1"),
  c("Maximum wind speed - 1", "Interaction - 1 - 2"),
  c("Maximum wind speed - 1", "Main effect - 1 - 2"),
  c("Minimum pressure - 1", "Interaction - 1 - 3"),
  c("Minimum pressure - 1", "Main effect - 1 - 3"),
  c("Normalized damage amount - 2", "Interaction - 2 - 1"),
  c("Normalized damage amount - 2", "Main effect - 2 - 1"),
  c("Maximum wind speed - 2", "Interaction - 2 - 2"),
  c("Maximum wind speed - 2", "Main effect - 2 - 2"),
  c("Minimum pressure - 2", "Interaction - 2 - 3"),
  c("Minimum pressure - 2", "Main effect - 2 - 3"),
  c("Normalized damage amount - 3", "Interaction - 3 - 1"),
  c("Normalized damage amount - 3", "Main effect - 3 - 1"),
  c("Maximum wind speed - 3", "Interaction - 3 - 2"),
  c("Maximum wind speed - 3", "Main effect - 3 - 2"),
  c("Minimum pressure - 3", "Interaction - 3 - 3"),
  c("Minimum pressure - 3", "Main effect - 3 - 3")
  )) %>%
  rename(from = V1, to = V2) %>%
  mutate(color = factor(c(2, 3, 8, 2, 8, 8, 3, rep(8, 5), 2,
                   rep(8, 5), 3, rep(8, 11))))
mygraph <- graph_from_data_frame(edges)
p4 <- ggraph(mygraph, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_diagonal(aes(color = color), 
                     show.legend = FALSE, width = 1) +
  geom_node_point(size = 4) +
  theme_void() +
  scale_edge_color_manual(values = c(2, 2, 8))
p4
```

.footnote[
1. Steegen, S., Tuerlinckx, F., Gelman, A., and Vanpaemel, W. (2016) _Increasing transparency through a multiverse analysis._ Perspectives on Psychological Science, 11(5), 702-712,  https://doi.org/10.1177/1745691616658637

]
---

.nav-left[

# Multiverse Analysis (Steegen et al. 2016)<sup>1</sup>

+   Performing and communicating all alternative analyses based on reasonable alternative paths
+   Improves **transparency** and **robustness** of the reported findings

]

.nav-right[

+   Jung et al. (2014)<sup>2</sup> reported a statistically significant result

+   Simonsohn et al. (2020)<sup>3</sup> considered 1,728 reasonable paths with the same data set

]


```{r out.width="45%", warning=FALSE, message=FALSE, dpi=144}
p4 + scale_edge_color_manual(values = c(2, 8, 8))
```

.footnote[
1. Steegen, S., Tuerlinckx, F., Gelman, A., and Vanpaemel, W. (2016) _Increasing transparency through a multiverse analysis._ Perspectives on Psychological Science, 11(5), 702-712,  https://doi.org/10.1177/1745691616658637

2. Jung, K., Shavitt, S., Viswanathan, M., and Hilbe, J.M. (2014) _Female hurricanes are deadlier than male hurricanes._ Proceedings of the National Academy of Sciences, 111(24), 8782-8787, https://doi.org/10.1073/pnas.1402786111

3. Simonsohn, U., Simmons, J.P., and Nelson, L.D. (2020) _Specification curve analysis._ Nature Human Behavior, 4 1208-1214, https://doi.org/10.1038/s41562-020-0912-z
]

---

.nav-left[

# Multiverse Analysis (Steegen et al. 2016)<sup>1</sup>

+   Performing and communicating all alternative analyses based on reasonable alternative paths
+   Improves **transparency** and **robustness** of the reported findings

]

.nav-right[

+   Jung et al. (2014)<sup>2</sup> reported a statistically significant result

+   Simonsohn et al. (2020)<sup>3</sup> considered 1,728 reasonable paths with the same data set

# _Multiverse analysis can get very complex and difficult to manage._
]


```{r out.width="45%", warning=FALSE, message=FALSE, dpi=144}
p4 + scale_edge_color_manual(values = c(2, 8, 8))
```

.footnote[
1. Steegen, S., Tuerlinckx, F., Gelman, A., and Vanpaemel, W. (2016) _Increasing transparency through a multiverse analysis._ Perspectives on Psychological Science, 11(5), 702-712,  https://doi.org/10.1177/1745691616658637

2. Jung, K., Shavitt, S., Viswanathan, M., and Hilbe, J.M. (2014) _Female hurricanes are deadlier than male hurricanes._ Proceedings of the National Academy of Sciences, 111(24), 8782-8787, https://doi.org/10.1073/pnas.1402786111

3. Simonsohn, U., Simmons, J.P., and Nelson, L.D. (2020) _Specification curve analysis._ Nature Human Behavior, 4 1208-1214, https://doi.org/10.1038/s41562-020-0912-z
]

---
layout:false
class:middle

# .selected[`mverse` aims to help students and analysts  
create, explore, and examine multiverse with ease]

.third[
## .hidden[Systematic Syntax]
]
.third[
## .hidden[Familiar Syntax]
]
.third[
## .hidden[Interactive Learning Modules]
]

---
class:middle

# .block[`mverse` aims to help students and analysts  
create, explore, and examine multiverse with ease]

.third[
## .accent1[Systematic Syntax]
]
.third[
## .accent3[Familiar Syntax]
]
.third[
## .accent2[Interactive Learning Modules]
]

---
.nav-left[

# .accent1[Systematic Syntax]

## Without `mverse`

+   Redundant codes
+   Prone to errors
]

.main-right.half[

```{r eval=FALSE, echo=TRUE}
# universe 1 (Michael) & 2
data_1 <- hdata %>%
  filter(! name %in% c("Katrina", "Audrey"))
fit_1 <- glm(
  deaths ~ damage * femininity, 
  data = data_1, family = "poisson")
# extract interaction
est_1 <- summary(fit_1)$coef[4,] 
# extract main effect
est_2 <- summary(fit_1)$coef[3,]

# universe 3 & 4
fit_2 <- glm(
  deaths ~ max_wind_speed * femininity, 
  data = data_1, family = "poisson")
# extract interaction
est_3 <- summary(fit_2)$coef[4,] 
# extract main effect
est_4 <- summary(fit_2)$coef[3,]

# universe 5 & 6
fit_3 <- glm(
  deaths ~ min_pressure * femininity, 
  data = data_1, family = "poisson")
# extract interaction
est_5 <- summary(fit_3)$coef[4,] 
# extract main effect
est_6 <- summary(fit_3)$coef[3,]


# universe 7 (Mary) & 8
data_2 <- hdata %>%
  filter(! name %in% c("Katrina"))
fit_4 <- glm(
  deaths ~ damage * femininity, 
  data = data_2, family = "poisson")
# extract interaction
est_7 <- summary(fit_4)$coef[4,] 
# extract main effect
est_8 <- summary(fit_4)$coef[3,]

# universe 9 & 10
fit_5 <- glm(
  deaths ~ max_wind_speed * femininity, 
  data = data_2, family = "poisson")
# extract interaction
est_9 <- summary(fit_5)$coef[4,] 
# extract main effect
est_10 <- summary(fit_5)$coef[3,]

# universe 11 & 12
fit_6 <- glm(
  deaths ~ min_pressure * femininity, 
  data = data_2, family = "poisson")
# extract interaction
est_11 <- summary(fit_6)$coef[4,] 
# extract main effect
est_12 <- summary(fit_6)$coef[3,]

# universe 13 & 14
data_3 <- hdata %>%
  filter(! name %in% c("Katrina", "Audrey", "Andrew"))
fit_7 <- glm(
  deaths ~ damage * femininity, 
  data = data_3, family = "poisson")
# extract interaction
est_13 <- summary(fit_7)$coef[4,] 
# extract main effect
est_14 <- summary(fit_7)$coef[3,]

# universe 15 & 16
fit_8 <- glm(
  deaths ~ max_wind_speed * femininity, 
  data = data_3, family = "poisson")
# extract interaction
est_15 <- summary(fit_8)$coef[4,] 
# extract main effect
est_16 <- summary(fit_8)$coef[3,]

# universe 17 & 18
fit_9 <- glm(
  deaths ~ min_pressure * femininity, 
  data = data_3, family = "poisson")
# extract interaction
est_17 <- summary(fit_9)$coef[4,] 
# extract main effect
est_18 <- summary(fit_9)$coef[3,]
```
]


---
.nav-left[

# Systematic Syntax

## With `mverse`

.accent1[
+   **Avoids redundant codes**
+   **Less prone to errors**
+   **Easier to reproduce**
]]

.main-right.half[

```{r eval=TRUE, echo=TRUE}
library(mverse)
# create decision points, or branches
hurricane_outliers <- filter_branch(
  ! name %in% c("Katrina", "Audrey"),
  ! name %in% c("Katrina"),
  ! name %in% c("Katrina", "Audrey", "Andrew")
)
hurricane_strength <- mutate_branch(
  damage, max_wind_speed, min_pressure
)
model <- formula_branch(
  deaths ~ hurricane_strength * femininity
)
distribution <- family_branch(
  poisson
)
# create a mverse object add branches
mv <- create_multiverse(hdata) %>%
  add_filter_branch(hurricane_outliers) %>%
  add_mutate_branch(hurricane_strength) %>%
  add_formula_branch(model) %>%
  add_family_branch(distribution) %>%
  glm_mverse() # fit a glm model
```
]

---
.nav-left[

# .accent3[Familiar Syntax]

]

.main-right.half[

```{r eval=FALSE, echo=TRUE}
library(mverse)
# create decision points, or branches
hurricane_outliers <- filter_branch( 
  ! name %in% c("Katrina", "Audrey"),
  ! name %in% c("Katrina"),
  ! name %in% c("Katrina", "Audrey", "Andrew")
)
hurricane_strength <- mutate_branch( 
  damage, max_wind_speed, min_pressure
)
model <- formula_branch( 
  deaths ~ hurricane_strength * femininity
)
distribution <- family_branch( 
  poisson
)
# create a mverse object add branches
mv <- create_multiverse(hdata) %>%
  add_filter_branch(hurricane_outliers) %>% 
  add_mutate_branch(hurricane_strength) %>% 
  add_formula_branch(model) %>%
  add_family_branch(distribution) %>%
  glm_mverse() # fit a glm model
```
]

---

.nav-left[

# Familiar Syntax

+   .accent3[**Familiar verbs from `dplyr`**]

]

.main-right.half[

```{r eval=FALSE, echo=TRUE}
library(mverse)
# create decision points, or branches
hurricane_outliers <- filter_branch( #<<
  ! name %in% c("Katrina", "Audrey"),
  ! name %in% c("Katrina"),
  ! name %in% c("Katrina", "Audrey", "Andrew")
)
hurricane_strength <- mutate_branch( #<<
  damage, max_wind_speed, min_pressure
)
model <- formula_branch( 
  deaths ~ hurricane_strength * femininity
)
distribution <- family_branch( 
  poisson
)
# create a mverse object add branches
mv <- create_multiverse(hdata) %>%
  add_filter_branch(hurricane_outliers) %>% #<<
  add_mutate_branch(hurricane_strength) %>% #<<
  add_formula_branch(model) %>%
  add_family_branch(distribution) %>%
  glm_mverse() # fit a glm model
```
]

---

.nav-left[

# Familiar Syntax

+   Familiar verbs from `dplyr`
+   .accent3[**Familiar syntax from base R**]

]

.main-right.half[

```{r eval=FALSE, echo=TRUE}
library(mverse)
# create decision points, or branches
hurricane_outliers <- filter_branch( 
  ! name %in% c("Katrina", "Audrey"),
  ! name %in% c("Katrina"),
  ! name %in% c("Katrina", "Audrey", "Andrew")
)
hurricane_strength <- mutate_branch( 
  damage, max_wind_speed, min_pressure
)
model <- formula_branch( #<<
  deaths ~ hurricane_strength * femininity
)
distribution <- family_branch(  #<<
  poisson
)
# create a mverse object add branches
mv <- create_multiverse(hdata) %>%
  add_filter_branch(hurricane_outliers) %>% 
  add_mutate_branch(hurricane_strength) %>% 
  add_formula_branch(model) %>% #<<
  add_family_branch(distribution) %>% #<<
  glm_mverse() # fit a glm model #<<
```
]

---

.nav-left[

# Familiar Syntax

+   Familiar verbs from `dplyr`
+   Familiar syntax from base R
+   .accent3[**Familiar `tibble` format for summary of results**]

]

.main-right[
```{r eval=FALSE, echo=TRUE}
summary(mv, conf.int = TRUE)
```
]
.tiny.two-third[

```{r eval=TRUE, echo=FALSE}
kable(summary(mv), conf.int = TRUE)
```

]

---
.nav-left[

# Familiar Syntax

+   Familiar verbs from `dplyr`
+   Familiar syntax from base R
+   .accent3[**Familiar `tibble` format for summary of results**]

]

.main-right[
```{r eval=FALSE, echo=TRUE}
summary(mv, conf.int = TRUE)  %>%
  filter(term %in% c("hurricane_strength:femininity", "femininity")) %>%
  select(hurricane_outliers_branch, hurricane_strength_branch, 
         term, estimate, conf.low, conf.high)
```
]
.tiny.two-third[

```{r}
summary(mv, conf.int = TRUE)  %>%
  filter(term %in% c("hurricane_strength:femininity", "femininity")) %>%
  select(hurricane_outliers_branch, hurricane_strength_branch, term, estimate, conf.low, conf.high) %>%
  kable()
```

]

---

.nav-left[

# Familiar Syntax

+   Familiar verbs from `dplyr`
+   Familiar syntax from base R
+   .accent3[**Familiar `tibble` format for summary of results**]

]

.main-right[

```{r out.width="100%", fig.asp=9/16, dpi=144}
ann <- paste0(
  "9 out of 18 universes in the example<br />",
  "found <b style='color:#DF536B;'>statistically siginificant results</b><br />",
  "while the other half found <b>no significant evidence</b><br />",
  "to support the hypothesis."
)
summary(mv, conf.int = TRUE)  %>%
  filter(term %in% c("hurricane_strength:femininity", "femininity")) %>%
  select(universe, term, estimate, conf.low, conf.high) %>%
  mutate(x = rank(estimate)) %>%
  ggplot(aes(x = x, y = exp(estimate), ymin = exp(conf.low), ymax = exp(conf.high),
             color = conf.low * conf.high > 0)) +
  theme_void() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_pointrange(show.legend = FALSE) +
  scale_color_manual(values = c(8, 2)) +
  annotate("richtext", label = ann, x = 0, y = 3.5, hjust = 0, vjust = 1,
           size = 6, color = NA, text.color = "#808080")
```

]

---
layout:false
class: dark, full
.nav-left[
# .accent2[Interactive Learning Modules]

+   Interactive modules to guide mutliverse analysis education
+   They do not require knowledge of coding in R


_The demo is a prototype and development is in progress_

]

.main-right[
<iframe src="https://micbon.shinyapps.io/mverse-ssc21-demo/" title="Interactive Presentation" style="border:none;width:90%;height:100%;"></iframe>
]

---
class:middle

.block[
.third[
## .accent1[Systematic Syntax]
]
.third[
## .accent3[Familiar Syntax]
]
.third[
## .accent2[Interactive Learning Modules]
]
]

# .selected[`mverse` can be used to introduce the idea of  
_multiverse_ to students at any level of their study]

---
layout:false
.nav-left[
# .selected[Development]
]
.main-right[
.timeline[
.item[
### In Development

Interactive learning modules are being developed

Visualizations for exploring and reporting multiverse analysis are also in development
]
.item[
### GitHub
      
`remote::install_github('mverseanalysis/mverse')` for the latest stable version

Currently supports multiverse analysis with `t.test`, `lm`, and `glm`
]
.item[
### CRAN

Plan to submit the first version in summer 2021 (no interactive learning modules)
]]]
---
class: full, middle, center

# Thank you
## .selected[
michael jongho moon  
June 7, 2021
]
